Taking Linear Logic Apart

### 2018-05-23 doc/abstract/main.tex
Put the \clearpage back in front of the bibliography
rewrite based on feedback

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
\usepackage{breakurl} % Not needed if you use pdflatex only.
\usepackage{underscore} % Only needed if you use pdflatex.
%\usepackage{parskip}
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{definition}[lemma]{Definition}
\input{preamble-typing}
\usepackage[sectionbib,sort&compress]{natbib}

@@

\begin{abstract}
  Process calculi based in logic, such as \piDILL and \cp, provide a foundation for deadlock-free concurrent programming. However, in previous work, there is a mismatch between the rules for constructing proofs and the term constructors of the \textpi-calculus. We introduce the Hypersequent Classical Processes (\hcp), which addresses this mismatch by using hypersequents (collections of sequents) to register parallelism in the typing judgements. We prove that \hcp enjoys deadlock-freedom and a series of properties that relate it back to \cp.
\end{abstract}

@@

There is no independent parallel term $\tm{(\piPar{P}{Q})}$ in the grammar of \cp terms. Instead, the parallel composition shows up in any term which corresponds to a binary rule. Even if we were to add an independent parallel composition via the \textsc{Mix}-rule, as suggested in the original presentation of \cp~\citep{wadler2012}, there would be no way to allow the composed process $\tm{P}$ and $\tm{Q}$ to communicate as in the standard \textpi-calculus, as there is no independent name restriction either! Instead, synchronisation is governed by the ``cut'' operator $\tm{\piNew{x}{(\piPar{P}{Q})}}$, which composes $\tm{P}$ and $\tm{Q}$, enabling them to communicate along $\tm{x}$. Worse, if we naively added an independent parallel composition as well as a name restriction, using the rules shown below, we would lose cut elimination, and therefore deadlock-freedom!
\begin{center}
  \begin{prooftree*}
    \AXC{$\seq[ P ]{ \Gamma }$}
    \AXC{$\seq[ Q ]{ \Delta }$}
    \NOM{Mix}
    \BIC{$\seq[ \piPar{P}{Q} ]{ \Gamma , \Delta }$}
  \end{prooftree*}
  \begin{prooftree*}
    \AXC{$\seq[ P ]{ \Gamma, \tmty{x}{A}, \tmty{y}{A^\bot} }$}
    \NOM{``Cut''}
    \UIC{$\seq[ \piNew{xy}{P} ]{ \Gamma }$}
  \end{prooftree*}
\end{center}

@@

In this paper, we introduce the Hypersequent Classical Processes (\hcp), which addresses this mismatch. The key insight is to register parallelism in the typing judgements using hypersequents~\citep{avron1991}, a technique from logic which generalises judgements from one sequent to many. This allows us to take apart the term constructs used in Classical Processes (\cp) to more closely match those of the standard \textpi-calculus. We proceed as follows. We start by introducing \cp (\cref{sec:cp}). Then, we introduce \hcp and prove it enjoys subject reduction and progress (\cref{sec:hcp}). We prove that every \cp process is an \hcp process, and that \hcp supports the same communication protocols as \cp, and no more (\cref{sec:cp2hcp}). Finally, we discuss related work (\cref{sec:related-work}).

@@

In this section, we introduce \cp. In order to keep the discussion of \hcp in \cref{sec:hcp} simple, we restrict ourselves to the multiplicative-applicative subset of \cp. We foresee no problems in extending the proofs in \cref{sec:hcp} to cover the remaining features of \cp (polymophism and the exponentials).

@@

The term language of \cp is a variant of the \textpi-calculus. The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names. The construct $\tm{\cpLink{x}{y}}$ links two channels~\citep{sangiorgi1996,boreale1998}, forwarding messages received on $\tm{x}$ to $\tm{y}$ and vice versa. The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel. Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$ and $\tm{Q}$. In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote input, square brackets denote output. We use bound output~\citep{sangiorgi1996}, meaning that both input and output bind a new name. In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$. In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$, while $\tm{x}$ is only bound in $\tm{Q}$.

@@

  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:cp-reduction}. The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of
  $\Longrightarrow$.

@@

\subsection{Types}
Channels in \cp are typed using a session type system which corresponds to classical linear logic. A process communicating over a channel of type \ty{A \tens B} must be ready to split into two independent processes (two processes which share no channels), ready to communicate over a channel of type \ty{A} and \ty{B}, respectively, and process communicating over a channel of type \ty{A \parr B} is dual to this, and therefore must be ready to start communicating concurrently with two independent processes. A process communicating over a channel of type \ty{A \plus B} can choose to communicate either as \ty{A} or \ty{B}, and a process communicating over a channel of type \ty{A \with B} must be able to communicate either as \ty{A} or \ty{B}.
\begin{definition}[Types]\label{def:cp-types}
  \[
    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}
    \end{array}
  \]
\end{definition}\noindent
Duality plays a crucial role in both linear logic and session types. In \cp, the two endpoints of a channel are assigned dual types. This ensures that, for instance, whenever a process \emph{sends} across a channel, the process on the other end of that channel is waiting to \emph{receive}. Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$. Duality is an involutive function.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
Environments associate channels with types. Names in environments must be unique, and environments $\ty{\Gamma}$ and $\ty{\Delta}$ can only be combined as $\ty{\Gamma}, \ty{\Delta}$ if $\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \ty{\emptycontext} \mid \ty{\Gamma}, \tmty{x}{A}
  \)
\end{definition}\noindent
\begin{figure}[htb]
  \begin{center}\cpInfAx    \cpInfCut   \end{center}
  \begin{center}\cpInfTens  \cpInfParr  \end{center}
  \begin{center}\cpInfBot   \cpInfOne   \end{center}
  \begin{center}\cpInfPlus1 \cpInfPlus2 \end{center}
  \begin{center}\cpInfWith              \end{center}
  \begin{center}\cpInfNil   \cpInfTop   \end{center}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\end{figure}\noindent
Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}\dots\tmty{x_n}{A_n}}$ denotes
  that the process \tm{P} communicates along channels $\tm{x_1}\dots\tm{x_n}$
  following protocols $\ty{A_1}\dots\ty{A_n}$.
  Typing judgements are derived using rules in \cref{fig:cp}.
\end{definition}\noindent

\subsection{Metatheory}
\cp enjoys subject reduction, progress, and termination~\citep{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}%
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$. If the last rule is \textsc{Cut}, there are four cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if either side of the cut is an axiom, we apply \cpRedAxCut1;
  \item
    if either side of the cut is itself a cut, we recursively eliminate the cut;
  \item
    if both sides are logical rules acting on the cut formula, we apply
    one of the $\beta$-rules;
  \item
    otherwise, at least one side is a logical rule acting on a formula other
    than the cut formula, in which case we apply one of the $\kappa$-rules.
  \end{enumerate*}
\end{proof}%
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts. However, each of these cuts is smaller, in the sense that the type of the channel on which the communication takes place is smaller. Each reduction eliminates a connective. Furthermore, each instance of the structural congruence preserves the size of the cut. Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}

@@

The term language of \hcp is a variant of \cp where the term constructs have been taken apart into \textpi-calculus primitives.

@@

  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under the rules in \cref{fig:hcp-reduction}. The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of $\Longrightarrow$.

@@

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar. Note that each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$ if you are also holding its other end-point''. This prevents self-locking processes, \eg $\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules in
  \cref{fig:hcp}.
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.
\end{definition}\noindent
\begin{figure}[htb]
  \begin{center}\hccpInfAx        \hccpInfCycle \end{center}
  \begin{center}\hccpInfMix       \hccpInfHalt  \end{center}
  \begin{center}\hccpInfBoundTens \hccpInfParr  \end{center}
  \begin{center}\hccpInfOne       \hccpInfBot   \end{center}
  \begin{center}\hccpInfPlus1     \hccpInfPlus2 \end{center}
  \begin{center}\hccpInfWith                    \end{center}
  \begin{center}\hccpInfNil       \hccpInfTop   \end{center}
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}

\subsection{Metatheory}
\hcp enjoys subject reduction, progress, and termination.
\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}%
\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}%
\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$. If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}%
\begin{theorem}[Termination]\label{thm:hcp-termination}
  If $\seq[P]{\mathcal{G}}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  As \cref{thm:cp-termination}.
\end{proof}

@@

  \label{fig:disentangle}

@@

%\clearpage
\bibliographystyle{apalike}
\bibliography{main}
\end{document}
%%% Local Variables:
%%% TeX-master: "main"
%%% End:

>>>>

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{Linearity and TLLA 2018}
\usepackage[british]{babel}
\usepackage[all]{foreign}
\usepackage[fleqn]{amsmath}
\usepackage{amsthm}
\allowdisplaybreaks
\usepackage[inline]{enumitem}
%\usepackage{breakurl} % Not needed if you use pdflatex only.
%\usepackage{underscore} % Only needed if you use pdflatex.
%\usepackage{parskip}
\usepackage{textcomp,textgreek,upgreek}
\usepackage{nameref,hyperref}
\usepackage[nameinlink,noabbrev,capitalize]{cleveref}
\hypersetup{hidelinks,final}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{definition}[lemma]{Definition}
\input{preamble-typing}
\usepackage[sectionbib,sort&compress]{natbib}

@@

\begin{abstract}
  Process calculi based on logic, such as \piDILL and \cp, provide a foundation for deadlock-free concurrent programming. However, in previous work, there is a mismatch between the rules for constructing proofs and the term constructors of the \textpi-calculus. We introduce Hypersequent Classical Processes (\hcp), which addresses this mismatch using hypersequents (collections of sequents) to register parallelism in the typing judgements. We prove that \hcp enjoys deadlock-freedom and a series of properties that relate it back to \cp.
\end{abstract}

@@

There is no independent parallel term $\tm{(\piPar{P}{Q})}$ in the grammar of \cp terms. Instead, parallel composition shows up in any term which corresponds to a typing rule which splits the context. Even if we were to add an independent parallel composition via the \textsc{Mix}-rule, as suggested in the original presentation of \cp~\citep{wadler2012}, there would be no way to allow the composed process $\tm{P}$ and $\tm{Q}$ to communicate as in the standard \textpi-calculus, as there is no independent name restriction either! Instead, synchronisation is governed by the ``cut'' operator $\tm{\piNew{x}{(\piPar{P}{Q})}}$, which composes $\tm{P}$ and $\tm{Q}$, enabling them to communicate along $\tm{x}$. Worse, if we naively add an independent parallel composition as well as a name restriction, using the rules shown below, we lose cut elimination, and therefore deadlock-freedom!
\begin{center}
  \begin{prooftree*}
    \AXC{$\seq[ P ]{ \Gamma }$}
    \AXC{$\seq[ Q ]{ \Delta }$}
    \NOM{Mix}
    \BIC{$\seq[ \piPar{P}{Q} ]{ \Gamma , \Delta }$}
  \end{prooftree*}
  \begin{prooftree*}
    \AXC{$\seq[ P ]{ \Gamma, \tmty{x}{A}, \tmty{y}{A^\bot} }$}
    \NOM{``Cut''}
    \UIC{$\seq[ \piNew{xy}{P} ]{ \Gamma }$}
  \end{prooftree*}
\end{center}

@@

In this paper, we introduce Hypersequent Classical Processes (\hcp), which addresses this mismatch. The key insight is to register parallelism in the typing judgements using hypersequents~\citep{avron1991}, a technique from logic which generalises judgements from one sequent to many. This allows us to take apart the term constructs used in Classical Processes (\cp) to more closely match those of the standard \textpi-calculus. We proceed as follows. We start by introducing \cp (\cref{sec:cp}). Then, we introduce \hcp and prove it enjoys subject reduction and progress (\cref{sec:hcp}). We prove that every \cp process is an \hcp process, and that \hcp supports the same communication protocols as \cp, and no more (\cref{sec:cp2hcp}). Finally, we discuss related work (\cref{sec:related-work}).

@@

In this section, we introduce \cp. In order to keep the discussion of \hcp in \cref{sec:hcp} simple, we restrict ourselves to the multiplicative-additive subset of \cp. We foresee no problems in extending the proofs in \cref{sec:hcp} to cover the remaining features of \cp (polymophism and the exponentials).

@@

The term language of \cp is a variant of the \textpi-calculus. The variables $\tm{x}$, $\tm{y}$, and $\tm{z}$ range over channel names. The construct $\tm{\cpLink{x}{y}}$ links two channels~\citep{sangiorgi1996,boreale1998}, forwarding messages received on $\tm{x}$ to $\tm{y}$ and vice versa. The construct $\tm{\cpCut{x}{P}{Q}}$ creates a new channel $\tm{x}$, and composes two processes, $\tm{P}$ and $\tm{Q}$, which communicate on $\tm{x}$, in parallel. Therefore, in $\tm{\cpCut{x}{P}{Q}}$ the name $\tm{x}$ is bound in both $\tm{P}$ and $\tm{Q}$. In $\tm{\cpRecv{x}{y}{P}}$ and $\tm{\cpSend{x}{y}{P}{Q}}$, round brackets denote input, square brackets denote output. \cp uses bound output~\citep{sangiorgi1996}, meaning that both input and output bind a new name. In $\tm{\cpRecv{x}{y}{P}}$ the new name $\tm{y}$ is bound in $\tm{P}$. In $\tm{\cpSend{x}{y}{P}{Q}}$, the new name $\tm{y}$ is only bound in $\tm{P}$, while $\tm{x}$ is only bound in $\tm{Q}$.

@@

  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under rules in \cref{fig:cp-reduction}. The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of $\Longrightarrow$. The relation $\Longrightarrow^{+}$ is the transitive closure of $\Longrightarrow$.

@@

\subsection{Types}
Channels in \cp are typed using a session type system which corresponds to classical linear logic.
\begin{definition}[Types]\label{def:cp-types}
  \[
    \begin{array}{rclrcl}
      \ty{A}, \ty{B}, \ty{C}
           :=& \ty{A \tens B} &\text{pair of independent processes}
      &  \mid& \ty{\one}      &\text{unit for} \; {\tens}
      \\ \mid& \ty{A \parr B} &\text{pair of interdependent processes}
      &  \mid& \ty{\bot}      &\text{unit for} \; {\parr}
      \\ \mid& \ty{A \plus B} &\text{internal choice}
      &  \mid& \ty{\nil}      &\text{unit for} \; {\plus}
      \\ \mid& \ty{A \with B} &\text{external choice}
      &  \mid& \ty{\top}      &\text{unit for} \; {\with}
    \end{array}
  \]
\end{definition}\noindent

A channel of type \ty{A \tens B} represents a pair of channels, which communicate with two independent processes---that is to say, two processes who share no channels. A process acting on a channel of type \ty{A \tens B} will send one endpoint of a fresh channel, and then split into a pair of independent processes. One of these processes will be responsible for an interaction of type \ty{A} over the fresh channel, while the other process continues to interact as \ty{B}.

A channel of type \ty{A \parr B} represents a pair of interdependent channels, which are used within a single process. A process acting on a channel of type \ty{A \parr B} will receive a channel to act on, and communicate on its channels in whatever order it pleases. This means that the usage of one channel can depend on that of another---\eg the interaction of type \ty{B} could depend on the result of the interaction of type \ty{A}, or vise versa, and if \ty{A} and \ty{B} are complex types, their interactions could likewise interweave in complex ways.

A process acting on a channel of type \ty{A \plus B} either sends the value \tm{\text{inl}} to select an interaction of type \ty{A} or the value \tm{\text{inr}} to select one of type \ty{B}. 
A process acting on a channel of type \ty{A \with B} receives such a value, and then offers an interaction of either type \ty{A} or \ty{B}, correspondingly.

Duality plays a crucial role in both linear logic and session types. In \cp, the two endpoints of a channel are assigned dual types. This ensures that, for instance, whenever a process \emph{sends} across a channel, the process on the other end of that channel is waiting to \emph{receive}. Each type $\ty{A}$ has a dual, written $\ty{A^\bot}$. Duality is an involutive function \ie $\ty{(A^\bot)^\bot} = \ty{A}$.
\begin{definition}[Duality]\label{def:cp-negation}
  \[
    \setlength{\arraycolsep}{3pt}
    \begin{array}{lclclclclclclcl}
               \ty{(A \tens B)^\bot} & = & \ty{A^\bot \parr B^\bot}
      &\qquad& \ty{\one^\bot}        & = & \ty{\bot}
      &\qquad& \ty{(A \parr B)^\bot} & = & \ty{A^\bot \tens B^\bot}
      &\qquad& \ty{\bot^\bot}        & = & \ty{\one}
      \\       \ty{(A \plus B)^\bot} & = & \ty{A^\bot \with B^\bot}
      &\qquad& \ty{\nil^\bot}        & = & \ty{\top}
      &\qquad& \ty{(A \with B)^\bot} & = & \ty{A^\bot \plus B^\bot}
      &\qquad& \ty{\top^\bot}        & = & \ty{\nil}
    \end{array}
  \]
\end{definition}\noindent
Environments associate channels with types. Names in environments must be unique, and environments $\ty{\Gamma}$ and $\ty{\Delta}$ can only be combined as $\ty{\Gamma, \Delta}$ if $\ty{\fv{\Gamma}} \cap \ty{\fv{\Delta}} = \varnothing$.
\begin{definition}[Environments]\label{def:cp-environments}
  \(
    \ty{\Gamma}, \ty{\Delta}, \ty{\Theta} := \ty{\emptycontext} \mid \ty{\Gamma}, \tmty{x}{A}
  \)
\end{definition}\noindent
\begin{figure}[htb]
  \paragraph*{Structural rules}
  \begin{center}\cpInfAx    \cpInfCut   \end{center}
  \paragraph*{Logical rules}
  \begin{center}\cpInfTens  \cpInfParr  \end{center}
  \begin{center}\cpInfBot   \cpInfOne   \end{center}
  \begin{center}\cpInfPlus1 \cpInfPlus2 \end{center}
  \begin{center}\cpInfWith              \end{center}
  \begin{center}\cpInfNil   \cpInfTop   \end{center}
  \caption{Classical Processes, typing.}
  \label{fig:cp}
\end{figure}\noindent
Typing judgements associate processes with collections of typed channels.
\begin{definition}[Typing judgements]\label{def:cp}
  A typing judgement $\seq[{ P }]{\tmty{x_1}{A_1}, \dots, \tmty{x_n}{A_n}}$ denotes that the process \tm{P} communicates along channels $\tm{x_1}$, \dots, $\tm{x_n}$ following protocols $\ty{A_1}$, \dots, $\ty{A_n}$. Typing judgements are derived using rules in \cref{fig:cp}.
\end{definition}\noindent

\subsection{Metatheory}
\cp enjoys subject reduction, termination, and progress~\citep{wadler2012}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}%
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts. However, each of these cuts is smaller, measured in the size of the cut term. Furthermore, each instance of the structural congruence preserves the size of the cut. Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\seq[P]{\Gamma}$. If the last rule is \textsc{Cut}, there are four cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if either side of the cut is an axiom, we apply \cpRedAxCut1;
  \item
    if either side of the cut is itself a cut, we recursively eliminate the cut;
  \item
    if both sides are logical rules acting on the cut formula, we apply
    one of the $\beta$-rules;
  \item
    otherwise, at least one side is a logical rule acting on a formula other
    than the cut formula, in which case we apply one of the $\kappa$-rules.
  \end{enumerate*}
\end{proof}%

@@

The term language of \hcp is a variant of \cp where the term constructs have been taken apart into primitives which more closely resemble the \textpi-calculus primitives.

@@

  Reductions are described by the smallest relation $\Longrightarrow$ on process terms closed under the rules in \cref{fig:hcp-reduction}. The relation $\Longrightarrow^\star$ is the reflexive, transitive closure of $\Longrightarrow$. The relation $\Longrightarrow^{+}$ is the transitive closure of $\Longrightarrow$.

@@

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar. Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$, which can be read as ``you cannot act on one end-point of $x$ if you are also holding its other end-point''. This prevents self-locking processes, \eg $\tm{\piBoundSend{x}{}{\piRecv{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules in
  \cref{fig:hcp}.
  Furthermore, each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.
\end{definition}\noindent
\begin{figure}[htb]
  \paragraph*{Structural rules}
  \begin{center}\hccpInfAx        \hccpInfCycle \end{center}
  \begin{center}\hccpInfMix       \hccpInfHalt  \end{center}
  \paragraph*{Logical rules}
  \begin{center}\hccpInfBoundTens \hccpInfParr  \end{center}
  \begin{center}\hccpInfOne       \hccpInfBot   \end{center}
  \begin{center}\hccpInfPlus1     \hccpInfPlus2 \end{center}
  \begin{center}\hccpInfWith                    \end{center}
  \begin{center}\hccpInfNil       \hccpInfTop   \end{center}
  \centering
  Each logical rule has the side condition that $\tm{x} \not\in \ty{\mathcal{G}}$.
  \caption{Hypersequent Classical Processes, typing.}
  \label{fig:hcp}
\end{figure}

\subsection{Metatheory}
\hcp enjoys subject reduction, termination, and progress.
\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}%
\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}%
\begin{theorem}[Termination]\label{thm:hcp-termination}
  If $\seq[P]{\mathcal{G}}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  As \cref{thm:cp-termination}.
\end{proof}
\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\mathcal{G}}$, then there exists a $\tm{Q}$ such that
  $\tm{P}\Longrightarrow^\star\tm{Q}$ and $\tm{Q}$ is not a cut or a mix.
\end{theorem}\vspace*{-0.75\baselineskip}%
\begin{proof}
  By induction on the derivation of $\seq[P]{\mathcal{G}}$. If the last rule is a cut, there are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the cut is the corresponding mix, we eliminate both as in
    \cref{thm:cp-progress};
  \item
    if the rule under the cut is an unrelated mix, we rewrite by \hccpEquivScopeExt{}, and recursively eliminate the cut;
  \item
    if the rule under the cut is a logical rule, we apply one of the $\kappa\nu$-rules.
  \end{enumerate*}
  If the last rule is mix, there are two cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if the rule under the mix is a cut or another mix, we recursively eliminate it;
  \item
    if the rule under the mix is a logical rule, we apply one of the
    $\kappa{\ppar}$-rules.
  \end{enumerate*}
\end{proof}%

@@

  \label{fig:hcp-disentangle}

@@

\clearpage
\bibliographystyle{apalike}
\bibliography{main}
\end{document}
%%% Local Variables:
%%% TeX-master: "main"
%%% End:
